Refugees in Toronto's Homeless Shelters: Evaluating the Need and Impact of Recent Policies and Increased Funding

### 2024-01-25 outputs/paper/paper.qmd
Added Abstract
Added the abstract to the paper
Added to Appendix
Moved the data cleaning portion into the appendix

Homelessness is a major problem in the city of Toronto, where over 9000 people are homeless. The causes of someone becoming homeless include loss of employment, domestic violence, substance abuse, mental health problems, and poor physical health. In Toronto, the drastic increases in house prices and rent costs have further worsened the situation, and around 77% of homeless people have been homeless for longer than six months [@citefred]. To help combat the problem, the City of Toronto offers numerous housing and homelessness services. For instance, they provide funds for shelter providers, which make up the majority of the shelters in Toronto, and also directly operate other shelters [@shelterinfo].

@@

The paper is structured as follows: Data, Results, and Discussion. The context behind the data used and the cleaning procedure is discussed in the data section. The results section introduces the graph and some summary statistics. The discussion section evaluates the trends observed and connects it to the underlying research topic, as well as discussing limitations of the analysis.

@@

The data set “Toronto Shelter System Flow” comprises of data from Shelter Management Information System (SMIS), which is used by Torontonian shelters to document each specific person who is entering and exiting the shelter system. Data is reported for every month for the past 5 years (2018 to 2023) for various populations, such as single adults, indigenous, youth, and refugees, and is refreshed every month. Moreover, a number of metrics for each population are reported, such as the number that moved to permanent housing, the number that have entered the system for the first time, the number who are below the age of 16, and numerous others [@dataset]. For the purpose of this paper, the focus is placed on refugees, thus, every other population was removed. The data set does not provide a specific metric on the total number of refugees in the system per month. However, data  split the total based on the age bracket they fell into ("ageunder16", "age16-24", "age25-44", ...) was included. Therefore, to calculate the total, a new column that sums all the “age” columns was added. After this calculation, every column, apart from the column on the number of people who entered the system for the first time, and the refugee population as a percentage of the total population, was removed (see @tbl-sample).  Given the use of the SMIS, no similar databases existed for the city of Toronto. @tbl-summary provides the summary statistics.

@@

## Sample of Cleaned Data
```{r}
#| echo: false
#| include: true
#| message: false
#| label: tbl-sample
#| tbl-cap: Sample of Cleaned Refugee Shelter Data

>>>>

Homelessness is a major problem in the City of Toronto, where over 9000 people are homeless. The causes of someone becoming homeless include loss of employment, domestic violence, substance abuse, mental health problems, and poor physical health. In Toronto, the drastic increases in house prices and rent costs have further worsened the situation, and around 77% of homeless people have been homeless for longer than six months [@citefred]. To help combat the problem, the City of Toronto offers numerous housing and homelessness services. For instance, they provide funds for shelter providers, which make up the majority of the shelters in Toronto, and also directly operate other shelters [@shelterinfo].

@@

The paper is structured as follows: Section 2 (Data), Section 3(Results), and Section 4 (Discussion). The context behind the data used is discussed in the data section. The results section introduces the graph and some summary statistics. The discussion section evaluates the trends observed and connects it to the underlying research topic, as well as discussing limitations of the analysis.

@@

The data set “Toronto Shelter System Flow” comprises of data from Shelter Management Information System (SMIS), which is used by Torontonian shelters to document each specific person who is entering and exiting the shelter system. Data is reported for every month for the past 5 years (2018 to 2023) for various populations, such as single adults, indigenous, youth, and refugees, and is refreshed every month. Moreover, a number of metrics for each population are reported, such as the number that moved to permanent housing, the number that have entered the system for the first time, the number who are below the age of 16, and numerous others [@dataset]. For the purpose of this paper, the focus is placed on refugees. The data cleaning process, and a sample of the data, is available in @sec-cleaning. Given the use of the SMIS, no similar databases existed for the city of Toronto. @tbl-summary provides the summary statistics.

@@

## Data Cleaning {#sec-cleaning}

Apart from refugees, every other population was removed. The data set does not provide a specific metric on the total number of refugees in the system per month. However, data  split the total based on the age bracket they fell into ("ageunder16", "age16-24", "age25-44", ...) was included. Therefore, to calculate the total, a new column that sums all the “age” columns was added. After this calculation, every column, apart from the column on the number of people who entered the system for the first time, and the refugee population as a percentage of the total population, was removed (see @tbl-sample).

### Sample of Cleaned Data
```{r}
#| echo: false
#| include: true
#| message: false
#| label: tbl-sample
#| tbl-cap: Sample of Cleaned Refugee Shelter Data
